Alone Together: Compositional Reasoning and Inference for Weak Isolation

### 2017-11-08 inference.tex
Formatting fixed.

$\lambda(\stl,\stg).~\stl \cup \F(\stg)$. Nonetheless, we keep
referring to $\F$ as state transformer. Since the execution is subject
to isolation-constrained interference, the inference judgment depends
on the isolation-constrained rely relation $\R$, which is used to
enforce the stability of the state transformer $\F$.  Recall that $\R$
is a tri-state rely relation over $\stl$, $\stg$ and $\stg'$, that
admits an interference from $\stg$ and $\stg'$ depending on the local

@@

(possible) prior state of the local database $\stl$, which depends on
the context (sequence of previous commands) of $c$, and computed by
the corresponding state transformer $\Fx$. Thus, the semantics of the
state transformer can be understood in terms of the RG judgment as
following (formalized as Theorem~\ref{thm:inference-sound} in

@@

    & & \hspace*{1.8in}}

@@

    & & \hspace*{1.8in}} {\bot}) \\

>>>>

  $\lambda(\stl,\stg).~\stl \cup \F(\stg)$. Nonetheless, for
  simplicity, we will continue to refer to $\F$ as state transformer.
  Since the execution is subject to isolation-constrained
  interference, the inference judgment depends on the
  isolation-constrained rely relation $\R$, which is used to enforce
  the stability of the state transformer $\F$.  Recall that $\R$ is a
  tri-state rely relation over $\stl$, $\stg$ and $\stg'$, that admits
  an interference from $\stg$ and $\stg'$ depending on the local

@@

  (possible) prior state of the local database $\stl$, which depends
  on the context (sequence of previous commands) of $c$, and computed
  by the corresponding state transformer $\Fx$. Thus, the semantics of
  the state transformer can be understood in terms of the RG judgment
  as following (formalized as Theorem~\ref{thm:inference-sound} in

@@

    & & \hspace*{1.5in}}

@@

    & & \hspace*{1.5in}} {\bot} \\

### 2017-11-08 introduction.tex
Formatting fixed.

Unfortunately, weak isolation admits behaviors that are difficult to
comprehend~\cite{berenson}. To quantify weak isolation anomalies,
Fekete \emph{et al.}~\cite{feketevldb09} devised and experimented with
a microbenchmark suite that executes transactions under a
weakly-isolated \emph{read committed} isolation level - the default
level for 8 of the 18 databases studied in~\cite{bailishotos}, and
found that 25 out of every 1000 rows in the database violate at least
one integrity constraint. Bailis \emph{et al.}~\cite{bailisferal} rely
on Rails' \emph{uniqueness validation} to maintain uniqueness of
records while serving Linkbench's~\cite{linkbench} insertion workload
(6400 records distributed over 1000 keys; 64 concurrent clients), and
report discovering more than 10 duplicate records.  Rails relies on
database transactions to validate uniqueness during insertions, which
is sensible if transactions are serializable, but incorrect under the
weak isolation level used in the experiments. The same study has found
that 13\% of all invariants among 67 open source Ruby-on-Rails
applications are at risk of being violated due to weak
isolation. Indeed, incidents of safety violations due to executing
applications in a weakly-isolated environment have been reported on
web services in production~\cite{starbucksbug, scimedbug}, including
in safety-critical applications such as bitcoin
exchanges~\cite{poloniexbug, bitcoinbug}. While enforcing
serializability for all transactions would be sufficient to avoid
these errors and anomalies, it would likely be an overly conservative
strategy; indeed, 75\% of the invariants studied in~\cite{bailisferal}
were shown to be preserved under some form of weak isolation.  When to
use weak isolation, and in what form, is therefore a prominent
question facing all database programmers.\footnote{This position has
been echoed by database researchers who lament the lack of a
better understanding of this problem; see e.g., {\tt
  http://www.bailis.org/blog/understanding-weak-isolation-is-a-serious-problem}.}

>>>>

Unfortunately, weak isolation admits behaviors that are difficult to
comprehend~\cite{berenson}. To quantify weak isolation anomalies,
Fekete \emph{et al.}~\cite{feketevldb09} devised and experimented with
a microbenchmark suite that executes transactions under a
weakly-isolated \emph{read committed} isolation level - the default
level for 8 of the 18 databases studied in~\cite{bailishotos}, and
found that 25 out of every 1000 rows in the database violate at least
one integrity constraint. Bailis \emph{et al.}~\cite{bailisferal} rely
on Rails' \emph{uniqueness validation} to maintain uniqueness of
records while serving Linkbench's~\cite{linkbench} insertion workload
(6400 records distributed over 1000 keys; 64 concurrent clients), and
report discovering more than 10 duplicate records.  Rails relies on
database transactions to validate uniqueness during insertions, which
is sensible if transactions are serializable, but incorrect under the
weak isolation level used in the experiments. The same study has found
that 13\% of all invariants among 67 open source Ruby-on-Rails
applications are at risk of being violated due to weak
isolation. Indeed, incidents of safety violations due to executing
applications in a weakly-isolated environment have been reported on
web services in production~\cite{starbucksbug, scimedbug}, including
in safety-critical applications such as bitcoin
exchanges~\cite{poloniexbug, bitcoinbug}. While enforcing
serializability for all transactions would be sufficient to avoid
these errors and anomalies, it would likely be an overly conservative
strategy; indeed, 75\% of the invariants studied in~\cite{bailisferal}
were shown to be preserved under some form of weak isolation.  When to
use weak isolation, and in what form, is therefore a prominent
question facing all database programmers.\footnote{This position has
been echoed by database researchers who lament the lack of a
better understanding of this problem; see e.g., \url{
  http://www.bailis.org/blog/understanding-weak-isolation-is-a-serious-problem}.}

### 2017-11-08 motivation.tex
Formatting fixed.

@@

\begin{figure}
\centering
\begin{ocaml}
let new_order (d_id, c_id, item_reqs) = atomically_do @@ fun () ->
  let dist = SQL.select1 District (fun d -> d.d_id = d_id) in
  let o_id = dist.d_next_o_id in
  begin
    SQL.update(* UPDATE *) District 
              (* SET *)(fun d -> {d with d_next_o_id = d.d_next_o_id + 1})
              (* WHERE *)(fun d -> d.d_id = d_id );
    SQL.insert(* INSERT INTO *) Order (* VALUES *){o_id=o_id;  
              o_d_id=d_id; o_c_id=c_id; o_ol_cnt=S.size item_reqs; };
    foreach item_reqs @@ fun item_req ->
      let stk = SQL.select1(* SELECT * FROM *) Stock 
                (* WHERE *)(fun s -> s.s_i_id = item_req.ol_i_id &&
                                     s.s_d_id = d_id)(* LIMIT 1 *) in
      let s_qty' = if stk.s_qty >= item_req.ol_qty + 10 
                  then stk.s_qty - item_req.ol_qty 
                  else stk.s_qty - item_req.ol_qty + 91 in
      SQL.update Stock (fun s -> {s with s_qty = s_qty'}) 
                       (fun s -> s.s_i_id = item_req.ol_i_id);
      SQL.insert Order_line {ol_o_id=o_id; ol_d_id=d_id; 
                             ol_i_id=item_req.ol_i_id; ol_qty=item_req.ol_qty}
  end

@@

This example does not exhibit any of the anomalies that
\emph{characterize} RC isolation~\cite{berenson}\footnote{Berenson et
al~\cite{berenson} characterize isolation levels in terms of the
anomalies they \emph{admit}. For example, RC is characterized by
\emph{lost writes} because it admits the anomaly.}. For instance,
there are no \emph{lost writes} since both concurrent transactions' writes
are present in the final state of the database.  Program analyses that
aim to determine appropriate isolation by checking for possible
manifestations of RC-induced anomalies would fail to identify grounds
for promoting the isolation level of \C{new\_order} to something
stronger.  Yet, if we take the semantics of the application into
account, it is quite clear that RC is not an appropriate isolation
level for \C{new\_order}.

>>>>

\begin{figure}[!t]
\centering
\begin{ocaml}
let new_order (d_id, c_id, item_reqs) = atomically_do @@ fun () ->
  let dist = SQL.select1 District (fun d -> d.d_id = d_id) in
  let o_id = dist.d_next_o_id in
  begin
    SQL.update(* UPDATE *) District 
              (* SET *)(fun d -> {d with d_next_o_id = d.d_next_o_id + 1})
              (* WHERE *)(fun d -> d.d_id = d_id );
    SQL.insert(* INSERT INTO *) Order (* VALUES *){o_id=o_id;  
              o_d_id=d_id; o_c_id=c_id; o_ol_cnt=S.size item_reqs; };
    foreach item_reqs @@ fun item_req ->
      let stk = SQL.select1(* SELECT * FROM *) Stock 
                (* WHERE *)(fun s -> s.s_i_id = item_req.ol_i_id &&
                                     s.s_d_id = d_id)(* LIMIT 1 *) in
      let s_qty' = if stk.s_qty >= item_req.ol_qty + 10 
                  then stk.s_qty - item_req.ol_qty 
                  else stk.s_qty - item_req.ol_qty + 91 in
      SQL.update Stock (fun s -> {s with s_qty = s_qty'}) 
                       (fun s -> s.s_i_id = item_req.ol_i_id);
      SQL.insert Order_line {ol_o_id=o_id; ol_d_id=d_id; 
                             ol_i_id=item_req.ol_i_id; ol_qty=item_req.ol_qty}
  end

@@

@@

This example does not exhibit any of the anomalies that
\emph{characterize} RC isolation~\cite{berenson}\footnote{Berenson et
al characterize isolation levels in terms of the
anomalies they \emph{admit}. For example, RC is characterized by
\emph{lost writes} because it admits the anomaly.}. For instance,
there are no \emph{lost writes} since both concurrent transactions' writes
are present in the final state of the database.  Program analyses that
aim to determine appropriate isolation by checking for possible
manifestations of RC-induced anomalies would fail to identify grounds
for promoting the isolation level of \C{new\_order} to something
stronger.  Yet, if we take the semantics of the application into
account, it is quite clear that RC is not an appropriate isolation
level for \C{new\_order}.
